---
title: "Data in Motion: Challenge 1 - Chipotle"
author: "Shazelle"
date: "2023-05-25"
output: pdf_document
---

# Scenario

You are a financial data analyst at Chipotle and your manager has tasked you with analyzing the most recent sales numbers. She has provided the following set of questions she would like answered.

## Load Packages

```{r}
library(tidyverse)
library(readr)
```



## Load Data

```{r}
url <- 'https://raw.githubusercontent.com/justmarkham/DAT8/master/data/chipotle.tsv'
chipotle_df <- read.csv(url, sep = '\t')
```



## View Data

```{r}
View(chipotle_df)
```

` `

Question 1. Which was the most-ordered item?

```{r}
#group the data by the item names and sum the total quantity
grouped_chipotle_df <- chipotle_df %>% group_by(item_name) %>% 
summarise(quantity = sum(quantity)) %>% arrange(-quantity)
```


```{r}
#return the top highest value
head(grouped_chipotle_df,1)
```

` `

Question 2. For the most-ordered item, how many items were ordered?

```{r}
max(grouped_chipotle_df$quantity)
```

` `

Question 3. What was the most ordered item in the choice_description column?

```{r}
#change all NULL values to NA
chipotle_df[chipotle_df=="NULL"] <- NA
```


```{r}
#omit all NA values during the grouping and count
choice_chipotle_df <- na.omit(chipotle_df) %>% group_by(choice_description) %>% 
summarise(total_count=n(),.groups = 'drop') %>% arrange(-total_count)
```


```{r}
#return top highest value
head(choice_chipotle_df,1)
```

` `

Question 4. How many items were ordered in total?

```{r}
sum(chipotle_df$quantity)
```

` `

Question 5. Turn the item price into a float.

```{r}
#omit the dollar sign then convert item price to float
chipotle_df$item_price <- as.double(gsub('[$,]','', chipotle_df$item_price))
str(chipotle_df)
```

` `

Question 6. How much was the revenue for the period in the dataset?

```{r}
sum(chipotle_df$item_price)
```

` `

Question 7. How many orders were made in the period?

```{r}
n_distinct(chipotle_df$order_id)
```

` `

Question 8. What is the average revenue amount per order?

```{r}
mean(chipotle_df$item_price)
```

` `

Question 9. How many different items are sold?
```{r}
count(grouped_chipotle_df)
```
